---
title: "q1"
output: html_document
bibliography: bibliography.ris
---


```{r setup-t1, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r deps-t1, echo=FALSE, include=FALSE, message=FALSE}
mirrorUrl = "http://cran.ma.imperial.ac.uk"

# Install and load all packages up-front!
if(!require(readxl)) install.packages("readxl", repos = mirrorUrl)
if(!require(Metrics)) install.packages("Metrics", repos = mirrorUrl)
if(!require(scales)) install.packages("scales", repos = mirrorUrl)
if(!require(data.table)) install.packages("data.table", repos = mirrorUrl)
#
library("readxl")
library("Metrics")
library("scales")
library("data.table")
```

<section>

#Task 1: Data Set Selection and Visualisation

##Premise

>You need to select a data set of your own choice (i.e. you may use a dataset already used before in the lab, or from the literature review) for the purposes of building training and validating the above type of classifiers 1-3. With the aid of R package visualise and justify the properties of the selected data set.

<!--
• Data Set summary of main properties 5
• Visualisation in R of main data set properties 5
-->

##Picking the data set

Firstly we need to load the data...

```{r read-excel-ww1}
#going to import the Excel spreadsheet WhiteWine dataset
wine.raw <- read_excel("../data/Whitewine.xlsx")
```

Here's a glance at the dataset

```{r show-wine-ww1}
head(wine.raw)
str(wine.raw)
```
</section>
