---
title: "q1"
output: html_document
bibliography: bibliography.ris
---


```{r setup-t5, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r deps-t5, echo=FALSE, include=FALSE, message=FALSE}
```

<section>

#Task 5: Build Train and Test a K-NN type Classifier

##Premise

>You need to construct, train and test K-NN type classifier in R. Train and test your K-NN classifier using the training and test sets generated based on the methods tried as part of the 2nd Task.

<!--
• Building of K-NN type classifier in R 4
• Training of K-NN type classifier in R 5
• Testing of K-NN type classifier in R 6
-->

... TODO Describe KNN a bit

<!--
model_knn &lt;- train(iris.training[, 1:4], iris.training[, 5], method=&#39;knn&#39;)
Note that making other models is extremely simple when you have gotten this far; You just
have to change the method argument.
Now that you have trained your model, it’s time to predict the labels of the test set that you
have just made and evaluate how the model has done on your data:
# Predict the labels of the test set
predictions&lt;-predict.train(object=model_knn,iris.test[,1:4], type=&quot;raw&quot;)
# Evaluate the predictions
table(predictions)
# Confusion matrix
confusionMatrix(predictions,iris.test[,5])

Additionally, you can try to do the
-->

##KNN modelling

###KNN with holdout

```{r knn-h_a}
knn.h_a <- train(dt.formula, data=holdout_age_groups.training, method = "knn")
```

```{r knn-h_a-summary}
plot(knn.h_a)
```

####KNN with holdout, testing

```{r knn-h_a_test_results}
knn.h_a_test <- predict(knn.h_a, holdout_age_groups.testing)
# summarize results
knn.h_a_test_cm <- confusionMatrix(knn.h_a_test, holdout_age_groups.testing$age_group)
knn.h_a_test_cm
```

###KNN with Repeated K-folds Cross-validation

```{r knn-r5kf8_a}
knn.r5kf8_a <- train(dt.formula, data=abalone_data_cleansed_age_groups, trControl = cv.train_control_8_5, method = "knn")
```

```{r knn-r5kf8_a-summary}
plot(knn.r5kf8_a)
```

####KNN with Repeated K-folds Cross-validation, testing

```{r knn-r5kf8_a_test_results}
knn.r5kf8_a_test <- predict(knn.r5kf8_a, holdout_age_groups.testing)
# summarize results
knn.r5kf8_a_test_cm <- confusionMatrix(knn.r5kf8_a_test, holdout_age_groups.testing$age_group)
knn.r5kf8_a_test_cm
```

###Naive Bayes with Leave-one-out Cross-validation

```{r knn-loo_a}
knn.loo_a <- train(dt.formula, data=abalone_data_cleansed_age_groups, trControl = loocv.train_control, method="knn")
```

```{r nb-loo_a-summary}
plot(knn.loo_a)
```

####Naive Bayes with Leave-one-out, training results

```{r nb-loo_a_test_results}
knn.loo_a_test <- predict(knn.loo_a, holdout_age_groups.testing)
# summarize results
knn.loo_a_test_cm <- confusionMatrix(knn.loo_a_test, holdout_age_groups.testing$age_group)
knn.loo_a_test_cm
```



###Accuracy comparison of KNN models

```{r}
knn.h_a_test_cm$overall[1]
knn.r5kf8_a_test_cm$overall[1]
knn.loo_a_test_cm$overall[1]
```


</section>
